Mapping mutational effects along the evolutionary landscape of HIV envelope

### 2017-12-12 paper/manuscript_Hugh.tex
Minor tweak to Hugh's edits

\hhcomment{I thought the original version of this paragraph was a bit redundant, so I have restructured it. Just an idea...}
In sum, we have shown how amino-acid preferences from two clade A Env variants can be compared to identify sites where mutational tolerance shifts with viral genotype, or between experiments and nature to identify sites under different pressure in the lab and in humans.
However, a challenge of studying HIV is that almost everything must be simplified to make it tractable.
Our studies are no different -- we are measuring viral replication in a T-cell line in the lab, not in infected humans.
Going forward, it will be interesting to determine how the results of experiments such as ours are affected by defined modifications in viral genotype or selective pressure.
For instance, although the Env variants we tested have 86\% protein identity, Env variants in nature can have as little as 65\% protein identity~\citep{lynch2009appreciating}.
Repeating our experiment with more divergent Env variants, including those from different clades, could quantify shifts in mutational effects over an even wider swath of Env's evolution.
Alternatively, adding additional selective pressures (e.g., antibody selection) to our experiment could help determine the extent that adaptive mutations differ in effects between Env variants.
Overall, such experiments should provide further insight into the forces that drive and constrain HIV's evolution.

>>>>

\hhcomment{I thought the original version of this paragraph was a bit redundant, so I have restructured it. Just an idea...}
In sum, we have shown how amino-acid preferences from two clade A Env variants can be compared to identify sites where mutational tolerance shifts with viral genotype, or between experiments and nature to identify sites under different pressure in the lab and in humans.
However, a challenge of studying HIV is that almost everything must be simplified to make it tractable.
Our studies are no different -- we are measuring viral replication in a T-cell line in the lab, not in infected humans.
Going forward, it will be interesting to determine how the results of experiments such as ours are affected by defined modifications in viral genotype or selective pressure.
For instance, although the Env variants we tested have 86\% protein identity, Env variants in nature can have as little as 65\% protein identity~\citep{lynch2009appreciating}.
Repeating our experiment with more divergent Env variants from transmitted-founder viruses, including those from different clades, could quantify shifts in mutational effects over an even wider swath of Env's evolution.
Alternatively, adding additional selective pressures (e.g., antibody selection) to our experiment could help determine the extent that adaptive mutations differ in effects between Env variants.
Overall, such experiments should provide further insight into the forces that drive and constrain HIV's evolution.
